---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r echo=T}


library(dplyr)
library(lattice)

activity <- read.csv(paste0("./repdata_data_activity/activity.csv"))

#convert date to date
activity$date <- as.Date(activity$date)

#convert steps to numeric
activity$steps <- as.numeric(activity$steps)

```




## What is mean total number of steps taken per day?

```{r echo=T}

hist(activity$steps[activity$steps != "NA"], main= "Histogram of Steps", xlab = "Number of Steps")

#Mean:
oldMean <- mean(activity$steps, na.rm = T)
print(oldMean)

#Median: 
oldMedian <- median(activity$steps, na.rm=T)
print(oldMedian)


```

## What is the average daily activity pattern?
```{r echo=T}


avg_steps <- activity %>% 
    group_by(interval) %>%
    summarize(steps=mean(steps, na.rm=T)) 

    plot(avg_steps$interval,avg_steps$steps, type="l", xlab="Interval", ylab="Steps")
    
#which 5-minute interval, on average, contains the maximum number of steps? 
    max_step_interval <- avg_steps$interval[avg_steps$steps==max(avg_steps$steps)]
    print(max_step_interval)
#Time of day is: 
    paste0(max_step_interval %/% 60, ":", max_step_interval%%60)

```


## Imputing missing values

```{r echo=T}

#Calculate and report the total number of missing values in the dataset
#(i.e. the total number of rows with NAs)
any(is.na(activity$date))
any(is.na(activity$interval))
# only steps has NAs
any(is.na(activity$steps))

#number of rows with NAs: 
length(activity[activity$steps=="NA",1])

#Devise a strategy for filling in all of the missing values in the dataset. The
#strategy does not need to be sophisticated. For example, you could use
#the mean/median for that day, or the mean for that 5-minute interval, etc.

# mean by 5-minute interval

activity <- activity %>%
    group_by(interval) %>% 
    mutate(steps_imp = mean(steps, na.rm=T))

#Create a new dataset that is equal to the original dataset but with the
#missing data filled in.

activity$steps <- ifelse(is.na(activity$steps), activity$steps_imp, activity$steps)

within(activity, rm(steps_imp))

# Make a histogram of the total number of steps taken each day and Calculate
# and report the mean and median total number of steps taken per day. 

hist(activity$steps, main= "Histogram of Steps", xlab = "Number of Steps")

#Mean
newMean <- mean(activity$steps)
print(newMean)

#Median
newMedian <- median(activity$steps)
print(newMedian)

```
Do these values differ from the estimates from the first part of the assignment?

No, the values are the same (`r newMean` = `r oldMean` and `r oldMedian` = `r newMedian`)


What is the impact of imputing missing data on the estimates of the total daily number of steps?

There is no impact on the mean or median. The histogram has higher counts due to the imputed values. 


## Are there differences in activity patterns between weekdays and weekends?

```{r echo=T}

#Create a new factor variable in the dataset with two levels – “weekday”
# and “weekend” indicating whether a given date is a weekday or weekend
# day.

activity <- activity %>% 
    mutate(day = weekdays(date)) %>%
    mutate(weekend = ifelse(day=="Saturday" | day=="Sunday", "weekend", "weekday"))
   
activity$weekend <- as.factor(activity$weekend)


# Make a panel plot containing a time series plot (i.e. type = "l") of the
# 5-minute interval (x-axis) and the average number of steps taken, averaged
# across all weekday days or weekend days (y-axis).

avg_steps_weekday <- activity %>% 
    group_by(weekend, interval) %>%
    summarize(steps=mean(steps))

xyplot(steps~interval | weekend, data=avg_steps_weekday, type="l", layout=c(1,2),
       xlab="Interval", ylab="Number of Steps")

```
